### 2015-04-22 thesis/thesis/chap4/chap4.tex
changes

\chapter{The Interaction between Probabilistic Commitments and Knowledge}\label{cha:PCTLKC}

In this chapter\footnote{The results of this chapter have been published in the journal of Expert Systems with Applications \cite{Sultan2014b}.}, we
put forward a method for capturing and verifying the interactions between the concepts of knowledge and social commitments in probabilistic MASs. The proposed method allows us to figure out the impact of knowledge and social commitments on each other in the presence of uncertainty. To express the two concepts simultaneously in systems exhibiting probabilistic behavior, we define a new modal logic called the probabilistic logic of knowledge and commitments (PCTL$^{kc}$) which combines two existing probabilistic logics namely, the probabilistic logic of knowledge (PCTLK) \cite{Wan2012,Wan2013} and the probabilistic logic of commitments (PCTLC) that has been introduced in Chapter \ref{cha:PCTLC}. In the current chapter, MASs are modeled using a new version of interpreted systems that captures the probabilistic behavior and accounts for the communication between interacting components. Based on the proposed logic, we introduce a new model checking procedure to check the compliance of target systems against some desirable properties -- with respect to knowledge and commitments-- written in PCTL$^{kc}$. %and report some verification results.

\section{Introduction} \label{Introduction}

The rapid increase of using software agents and MASs nowadays has led to the increasing demand of finding principled techniques for modeling and verifying such systems. Generally, to build effective open MASs, several aspects which have direct influence on the efficiency and effectiveness of
the entire system must be taken into account \cite{Konur2013}.
Among other aspects, knowledge and social commitments are of a great
interest in MASs. Social commitments have been a vital approach in agent societies to capture the communication between interacting agents for more than a decade. On the other hand, knowledge has been addressed in distributed systems since 1960s \cite{Wan2013}. Recently, Al-Saqqar et al. \cite{Al-Saqqar2014a} have demonstrated that these two concepts are closely interacting with each other in various real life scenarios.

Despite the large amount of work that has been done to model and represent various aspect of probabilistic MAS, none of the existing approaches addresses the concepts of knowledge and social commitments simultaneously. In fact, the problem of reasoning about and verifying the interaction between knowledge and social commitments in the presence of uncertainty has not been investigated yet. Interpreted systems formalism \cite{Fagin1995}
and Partially Observable Markov Decision Processes POMDPs (a
variant of MDP) are the most prominent traditions in the area of
modeling and representing stochastic MASs. These models are used
to traditionally interpret some logics defined to specify and
reason about some given properties of MASs. On the one hand,
interpreted systems formalism provides a natural and yet efficient
way for modeling MASs at different levels of abstractions (i.e.,
local and global). It has been extended in \cite{Halpern2003} and
further in \cite{Wan2012,Wan2013} to capture the probabilistic
behavior of epistemic MASs. Recently, it has been extended in
\cite{Bentahar2012} and \cite{El-Menshawy2013a} to account for the
communication that occur between interacting parties in
conventional MASs. The distinct point of the extended versions of
this formalism is that knowledge and commitments can be captured
through the use of what is called accessibility relations. The
accessibility relation for knowledge denotes the existence of
equivalent states for a given agent. That is, states where the
agent cannot distinguish between being in which one of them. For
commitments, accessibility relations capture the existence of
communication channel between the communicating agents and the
transferring of information from the sender to the receiver. On
the other hand, POMDPs have been widely used to model the
uncertainty of knowledge and behavior for stochastic agents
\cite{Huang2011}. An important point of POMDPs is that there is no distinction drawn between actions taken to change the state of the world and actions taken to gain information \cite{Kaelbling1996}. This is important
because, in general, every action has both types of effect.
However, solving these models comes at a very high computational
cost \cite{Melo2011}. In this chapter, we aim to examine the use of interpreted systems formalism to capture not only knowledge and commitments
independently, but also the interactions (combinations) of the two
aspects in stochastic systems. We also intent to verify these
interactions by means of model checking.

In terms of computational logics, most current proposals address each of knowledge and commitments in MASs independently (see for example
\cite{Baldoni2010,Bentahar2012,Delgado2009,El-Menshawy2013a,Giordano2007,Halpern2003,Huang2011,Lomuscio2007,Pham2007,Wan2013}).
However, in so many real world settings, these two concepts need to
interact with each other in order to ensure rich modeling at local
(agent) and global (MAS) levels. Nevertheless, it is a challenge to guarantee the correctness of the system's behavior due to the complex nature of the autonomous and heterogenous agents, especially when they have probabilistic characteristics \cite{Song2012}.

Applying model checking techniques that were originally introduced
for standard logics, such as LTL \cite{Pnueli1977}, CTL
\cite{Emerson1990}, or PCTL \cite{Hansson1994}, to the
verification of the interaction between knowledge and social
commitments in presence of uncertainty is not straightforward as
non of these logics can capture and express the relationship
between knowledge and social commitments in probabilistic
settings. In this chapter, we introduce a model checking technique to address this open issue.

The motivation for the incorporation of knowledge and commitments
in a probabilistic logic is provided by the fact that these two
concepts not only have an impact on each other, but also their
interaction is crucial in various real scenarios. For instance, in
the field of mobile applications, which are complex in nature,
there exist situations when accounting for the interaction between
knowledge and commitments improves the output of such
applications. Let us consider a simple scenario where receiver and
sender agents share an agreement, in which the receiver agrees to
pay the sender in return of the delivery of a service he has
requested. This can be represented as a commitment, in which the
receiver will be committed to the sender to pay once the service
is made available for him. Now, if everything goes well and the
receiver successfully makes his payment, the sender has to know
that the payment is made so that he does not ask the receiver to
pay again. Moreover, the receiver (who made the payment) has to
know that he has fulfilled his commitment to avoid making multiple
payments, and so on. However, those interactions are stochastic.
For instance, the commitment to pay is not going to be surely
satisfied.

To effectively specify such properties in the face of uncertainty,
the need for a logical tool that can express probabilistic
knowledge and commitments simultaneously is indeed confirmed.
Rather than building a logic from scratch to address the
underlying aspects, we combine logics dealing with these two
individual units in a single logic. We advocate the approach of
combining existing logics because it ensures the preservation of
important properties of the logics being combined
\cite{Konur2013}. In particular, we use the independent join (or
fusion) technique \cite{Franceschet2004}. Given two logics $\mathbb{A}$ and $\mathbb{B}$, we combine them in a new logic $\mathds{A} \oplus \mathds{B}$ which extends the expressive power of each one. In our case, suppose $\mathds{A}$ addresses probabilistic epistemic properties of
agents and $\mathds{B}$ addresses the social aspects (i.e.,
probabilistic commitments and their fulfilments) between
interacting agents. Their combination should be able to not only
express epistemic and social properties, but also express the
interaction between the two concepts (i.e., express them in a
single formula). Once the new combined logic is defined, we use
the PRISM tool \cite{Kwiatkowska2002} as the formal verification tool to verify it after its reduction to PCTL, the probabilistic branching-time logic \cite{Hansson1994}.

The contributions of this chapter are threefold. First, we present a new probabilistic version of interpreted systems to model MASs using the dimensions of knowledge and social commitments. The developed version merges two extended versions of the original formalism of interpreted systems introduced by Fagin and his colleagues \cite{Fagin1995}. Those versions are introduced respectively by 1) Halpern \cite{Halpern2003} and extended later
by Wan et al. \cite{Wan2012,Wan2013} to capture the stochastic
behavior of the system; and 2) by Bentahar et al.
\cite{Bentahar2012} and El-Menshawy et al. \cite{El-Menshawy2013a}
to model the communication between interacting parties. Second, we
introduce a new logic called Probabilistic Logic of Knowledge and
Commitment (PCTL$^{kc}$) to be able to capture and reason about
the interaction between knowledge and social commitments. The
logic we define combines the Probabilistic Computation Tree Logic
of Knowledge PCTLK \cite{Wan2012,Wan2013} and the Probabilistic
Computation Tree Logic of Commitments PCTLC \cite{Sultan2013}.
PCTLK and PCTLC are, in turn, extensions of the Probabilistic
Computation Tree Logic PCTL \cite{Hansson1994} with an epistemic
modality for the knowledge and a social modality for the
commitments and their fulfilments respectively. Third, we
introduce a new model checking technique to verify the proposed
logic (PCTL$^{kc}$). The introduced technique is a reduction-based
in which the problem of model checking PCTL$^{kc}$ is transformed
into the problem of model checking an existing logic called PCTL.
To achieve this reduction, new rules have been laid down to
transform the models of PCTL$^{kc}$ to MDPs to be suitable for
the PRISM model checker. We also devise some other rules to reduce
each PCTL$^{kc}$ formula into PCTL formula. By so doing, we can
build on the existing PRISM model checker by automating our
translation to verify some given properties written originally in
our new logic PCTL$^{kc}$. Figure \ref{approach-cha4} gives an overview of the proposed approach.

\begin{figure}[h]
\begin{center}
\includegraphics [width=14cm, height=5cm]{chap4/img/approach-cha4.eps}
  \caption{An approach for the interaction between knowledge and commitments}
\label{approach-cha4}
\end{center}
\end{figure}

The work presented in this chapter represents a new trend in the direction of capturing interactions between various aspects in MASs. It can be
seen as a first attempt to combine the notions of probability,
knowledge, and commitments in a single tool giving a new
expressive power ---in terms of expressing the individual aspects
as well as their combinations in the presence of uncertainty---,
and is therefore subject to new intuitions.

\section{The Probabilistic Logic of Knowledge and Commitment (PCTL$^{kc})$} \label{sec:PCTLKC-logic}

In this section, we introduce our new probabilistic logic of knowledge and commitment (PCTL$^{kc})$. The modal logic we introduce can express knowledge and social commitments simultaneously in the presence of uncertainty. It combines two existing probabilistic logics namely, the probabilistic logic of knowledge PCTLK \cite{Wan2012,Wan2013} and the probabilistic logic of commitments PCTLC \cite{Sultan2014a,Sultan2013}. We first present the syntax of our new logic, and then we define its semantics. We also define a new version of the probabilistic interpreted systems formalism over which the semantics of PCTL$^{kc}$ can be interpreted.

As we said earlier, the new logic PCTL$^{kc}$ contains a knowledge modality that doesn't exist in the logic defined previously in Chapter \ref{cha:PCTLC}. Therefore, we need first to define the model of PCTL$^{kc}$. In fact, the PCTL$^{kc}$ model is generated from an extended version of probabilistic interpreted systems \cite{Halpern2003,Wan2013} enriched by the social accessibility relations introduced in \cite{Bentahar2012,El-Menshawy2013a} as discussed in Chapter \ref{cha:background}.

\begin{definition}[Models]\label{def:models-pctlkc}

Given a set of atomic propositions $\Phi_p = (p,q,r, \ldots)$ and a set of agents $\texttt{Agt}=\{1,\ldots,n\}$,  the model $\mathfrak{M_2}=(S,\textbf{P},I,\sim_1, \ldots
,\sim_n,\{\sim_{i \rightarrow j}\}_{{(i,j)}\in \texttt{Agt}^2},\nu)$ is a tuple where:
%
\begin{itemize}
\item  $S \subseteq L_1 \times \ldots \times L_n$ is a countable
set of all reachable global states for the system. A state $s$ is
reachable iff there exists a sequence of transitions from an
initial state to $s$ in which the probability of each transition
is greater than $0$.

\item  $I \in S$ is an initial global state for the system.

\item  $\textbf{P}:S\times S\rightarrow [0,1]$ is a total
transition probability function defined as $\textbf{P}(s,
s')=\tau(s,a^{s \rightarrow s'}, s')$ iff there exists a joint
action $a=(a_1,\ldots,a_n) \in ACT$ such that\\
 $\sum_{i \in \texttt{Agt}} \tau_i(l_i(s),a^{l_i(s)\rightarrow l_i(s')},l_i(s')) > 0$ and $\sum_{s' \in S} \textbf{P}(s,s') =1$
for all $s \in S$.

\item  $\sim_i \subseteq S \times S$ is the epistemic
  accessibility relation for the agent $i$, such that for two global states $s$ and $s'$, we have: $s \sim_i s'$ iff
  $l_i(s)=l_i(s')$.

\item For each pair $(i,j) \in \texttt{Agt}^2$,
$\sim_{i\rightarrow j} \subseteq S \times S$ is a serial social
accessibility relation. $s \sim_{i\rightarrow j} s'$ is defined by
the following conditions:
      %
  \begin{enumerate}
       \item $l_i(s)=l_i(s')$.
       \item $Var_i \cap Var_j \neq \emptyset$ such that $\forall x \in Var_i \cap Var_j$ we have $l^{x}_i(s)\!=\!l^{x}_j(s')$.
       \item $\forall y \in Var_j\!-\! Var_i$ we have $l^{y}_j(s)\!=\!l^{y}_j(s')$.
  \end{enumerate}
       %
  \item  $\nu : S\rightarrow 2^{\Phi_p}$ is a function valuating states with atomic propositions.
\end{itemize}
\end{definition}
%

The difference between our new model $\mathfrak{M_2}$ and the model $\mathfrak{M_1}$ that has been proposed in Chapter \ref{cha:PCTLC} is that $\mathfrak{M_2}$ has the ability to model agents' knowledge in the system --in addition to modeling the commitment-based communication among interacting parties-- thanks to the epistemic accessibility relations that are integrated in the model. Technically, $\mathfrak{M_2}$ in an extended version of $\mathfrak{M_1}$ with epistemic accessibility relations. Computation paths of $\mathfrak{M_2}$ and probability space are defined as in Chapter \ref{cha:PCTLC}. %(Section \ref{sec: semantics of PCTLC}).

\subsection{Syntax of PCTL$^{kc}$} \label{sec:syntax-PCTLKC}
The logic we introduce in this section can be seen as an extension to the logic presented in Chapter \ref{cha:PCTLC} by adding an epistemic operator to PCTLC \cite{Sultan2014a,Sultan2013}. The resulting logic, i.e., PCTL$^{kc}$, will have the power to not only express the individual aspects of knowledge and social commitments in independent formulae, but also express combinations of the two concepts in the same formulae.

\begin{definition}[Syntax]\label{def:PCTLKC syntax}
Given a set of atomic propositions $\Phi_p$. Let $\verb"Agt"=\{1, \dots, n\}$ be a set of agents. The PCTL$^{kc}$ formulae are defined by the following BNF grammar:
%
\begin{align*}
    \varphi & ::= p~|~\neg \varphi~|~\varphi \vee \varphi~|~\mathcal{K}~|~\mathcal{C}~|~ \mathbb{P}_{\bowtie k} (\psi)~|~\mathbb{P}_{\bowtie k}(\mathcal{K})|~\mathbb{P}_{\bowtie k}(\mathcal{C})\\
    \psi & ::=\bigcirc \varphi ~ | ~ \varphi ~U~ \varphi~|~ \varphi~ U^{\leq m} ~ \varphi \\
    \mathcal{K} & ::= K_i \varphi \\
    \mathcal{C} & ::= C_{i\rightarrow j}\varphi ~| ~ Fu(C_{i\rightarrow j}\varphi)
\end{align*}
%
where $p\in\Phi_p$ is an atomic proposition and $\mathbb{P}_{\bowtie
k}$ is a probabilistic operator where $\bowtie \in\{<,\leq,>,\ge\}$ and $k\in [0,1]$ is a probability bound or threshold. $m \in\mathbb{N}^+ $ is a positive integer number reflecting the maximum number of transitions needed to reach a certain state. $\varphi$ and $\psi$ are state and path formulae interpreted over the states and paths of $\mathfrak{M_2}$ respectively. The Boolean connectives $\neg$ and $\vee$ are defined in the usual
way. Formulae $\mathcal{K}$ are state formulae called knowledge (epistemic) formulae and used to express the epistemic properties through the K$_i$ operator which stands for agent i knows. Formulae $\mathcal{C}$, called social formulae, are special state formulae in PCTL$^{kc}$ that can express social properties using the modal connectives $C_{i\rightarrow j}$ and $Fu(C_{i\rightarrow j})$ standing for ``commitment'' and ``fulfillment of commitment'' respectively. $\bigcirc, U$ and $U^{\leq m}$ stand for ``next time'', ``until'' and ``bounded until'' path modal connectives
respectively.

\end{definition}

\subsection{Semantics of PCTL$^{kc}$} \label{def:semantics-PCTLKC}

Given a model $\mathfrak{M_2}=(S,\textbf{P},I,\sim_1, \ldots
,\sim_n,\{\sim_{i \rightarrow j}\}_{{(i,j)}\in \texttt{Agt}^2},\nu)$, then $(\mathfrak{M_2},s) \models \varphi$ states that ``a state $s$ in the model $\mathfrak{M_2}$ satisfies a state formula $\varphi$, $(\mathfrak{M_2},\pi) \models \psi$ means that ``a path $\pi$ in the model $\mathfrak{M_2}$ satisfies a path formula $\psi$, and
$(\mathfrak{M_2},s) \models \mathbb{P}_{\bowtie k}(\psi)$ means that
``a state $s$ in $\mathfrak{M_2}$ satisfies $\mathbb{P}_{\bowtie k}(\psi)$ if the probability of taking a path from $s$ that satisfies $\psi$ is in the interval specified by $\bowtie k$''. When the model $\mathfrak{M_2}$ is clear from the context, we simply write the satisfaction relation $\models$ as follows: $s \models \varphi$ and $\pi \models \psi$. Furthermore,
for a given pair $(i,j)\in \texttt{Agt}^2$ of agents, we denote
the number of socially accessible states $s'$ from a given state $s$ such
that $s\sim_{i\rightarrow j}s'$ by $|s\sim_{i\rightarrow j}s'|$.
We also denote the number of epistemically accessible states $s'$ from a given state $s$ such that $s\sim_i s'$ by $|s\sim_i s'|$.

Finally, we define $|s\models \varphi|$ as follows:

\begin{center}
$|s\models \varphi|=
\begin{cases}
1,~~~\textrm{if}~ s\models\varphi\\
0,~~~\textrm{otherwise.}
\end{cases}$
\end{center}

\begin{definition}[\textbf{Satisfaction}]\label{def:semantics-pctlkc} Satisfaction of a PCTL$^{kc}$ formula in the model $\mathfrak{M_2}$ is recursively defined as follows:

%\noindent In what follows, we define the semantics of PCTL$^{kc}$ formulae.

>>>>

\chapter{Training Chapter}\label{cha:PCTLKC}

\section{Introduction}

Over the past years, online services have become an important part of many scalable business applications. The increasing reliance on online service providers has significantly influenced the way web services are engineered. Therefore, given the dynamic and unpredictable nature of the Internet, delivering high quality services is still a critical and challenging issue. One practical solution towards delivering high quality services is utilizing intelligent decision making agents. These agents aim at maximizing their gain by exploring the best ways to provide services that satisfy end users\cite{Zeng:2003:QDW:775152.775211, 10.1109/ARES.2008.7, Demirkan2013412, journals/tsc/ZhengZYB13, Josang:2007:STR:1225318.1225716}. However, agent-based web services are functionally limited in the sense that they cannot handle a large number of requests at the same time without compromising the quality of service provided. Recent developments have attempted to shift web services from passive models, consisting of individual components, to models made up of autonomous and group-based components that share common goals. In group-based models, interaction, composition, and cooperation are the key challenges \cite{ICWS2011-1, SCC2011-1, journals/mags/BaldoniBM10, journals/jcss/CasadoYT13} that directly impact the group's overall performance in achieving common goals. To that end, we see the emergence of web service \emph{communities}, which consist of services with similar functionalities but distinct nonfunctional properties \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. A \emph{community} of web services runs continuous performance assessment functions that regulate web services' interactions and manage their composition and cooperation.

The advantages of community of web services are that the community facilitates the discovery of web services and provides better quality of service compared to passive models with individual components. Communities act as abstract web services, communicating with external entities via the same standard protocols that a normal web service employs. But, they regulate the service process via sophisticated internal communication protocols, thereby providing services based on the combined work of a number of web services. The downside to communities is the complexity of management involved in finding and inviting individual web services and managing the overall quality of the combined work of several web services.
% because although they have similar functionality, they have different attitudes.
When interacting with communities of web services, users send their requests to the coordinators of the community, which play the role of community representative or access point. The community coordinator is responsible of receiving tasks and delivering services. Moreover, as community representative, it verifies the credentials of new web services before accepting them into the community and kicks web services that could harm the value of the community.

\textbf{Challenge and Problem Statement.} In recent work, communities of web services have been proposed in order to facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities. However, there is a twofold challenge that yet to be addressed. The first part is community development, which involves finding the right web services for the community, satisfying the objectives of all the web services and community coordinators involved within the community; and the second part is from the individual web service's side because each web service aims to join the community that would guarantee the best outcome. The advocated solutions \cite{conf/webist/MaamarLBTS07, 10.1109/ARES.2008.7, journals/soca/XuYLZB11, 10.1109/TSC.2012.12, managing-hela-jalel, DBLP:conf/IEEEscc/KhosravifarABT11, DBLP:conf/IEEEscc/LimTMB12} have attempted to address these research problems. These research problems are reflected by the two challenge parts:
\begin{enumerate}
	\item The solutions consider the architecture of centralized management for communities where most of the decisions are made by the centralized coordinator. The problem is that in real world scenarios, decisions made by independent service providers are highly distributed.
	\item The solutions either propose complex algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to find the optimal strategy to follow, or oversimplify the problem by eliminating important parameters and using approximation techniques to make the algorithms tractable \cite{10.1109/TSC.2012.12}.
\end{enumerate}
These approximation methods sometimes negatively influence the outcome because simplifying the constraints may cause important aspects of the problem to be ignored.

For example, instead of calculating the gain distribution using the complex shapely-value method, \cite{10.1109/TSC.2012.12} proposes a simple egalitarian way of distributing gain, which completely ignores the gain generated from collaborative work of sub-communities. Other categories of related work exist that do not focus on all the parties involved within the community \cite{DBLP:conf/IEEEscc/KhosravifarABT11}.
%And some other work which do not focus on rationality on web services or communities involved [refrences].
In \cite{journal-community-formation} we proposed a cooperative game theory based model for aggregating web services in communities. A centralized decision maker in communities based on a complete knowledge of available web service quality metrics and performance, has been used to form optimal and stable communities that maximize individual and group income. However, centrality and complete information are strong assumptions, which are not very compatible with real business scenarios.

\textbf{Contributions.} In this paper, we introduce DDM, a Distributed Decision Making model for  community formation that regulates web service agents’ decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. The decision model in extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and how to select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree is created from the past data and reflects comprehensive vision about agents' attitude of action selection. However, the tree is subject to get continuously modified to adapt with environmental changes.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems.
The decision model provides web services with enough information which helps those services to efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in decision making process and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. In other words, the web services agents are equipped with a recommender system to suggest beneficial action. Moreover, communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact their short-term and long-term goals; therefore, opting for best decision available.

To effectively generate the decision model for web services, we used real data set to extract web services' individual characteristics and used them to measure outcomes when they cooperate with one another. The data set has been extracted by real-world QoS evaluation results from 142 users on 4,532 Web services on 64 different time slots. Combining these data based on each web service point of view on different time slots, we acquired 5 different unique features for 4,532 web services. By engineering and extracting these features, we gathered functional and cooperative features for both individual web services and communities in different time slots. We were able to investigate the path a web service might take to achieve the best utility out of effective interaction with others. All paths and the outcomes are labeled to be utilized in the training model. Using cross validation sets, web services are able to compute the optimal hypothesis function (using logistic regression) that can be used to predict outcomes of cooperative work with other individual web services or communities. Our findings show that web services equipped with DDM have by far better outcomes in the form of cooperative service providers than the ones that either do not cooperate or randomly find communities to join.

\section{Preliminaries and Challenging Issues}\label{s:preliminaries}
%In this section, we first present the architecture of DDM. We explore the characteristics of intelligent service agents and the features we extract for training. To do this, we first discuss some preliminaries.
In this section, we discuss the preliminary concepts of communities of web service and we introduce the challenges of community formation.

\subsection{Web Services}\label{s:ws}

In recent years, online services have become a standard part of daily life around the globe. Many modern applications rely on web services from different providers. For instance, many mobile and tablet applications that have limited storage and processing power are merely aggregating information from different online services. Examples are vast, including weather forecasting, ticket selling, shopping apps, local maps and location services.

The World Wide Web Consortium (W3C) defines web services as ``software systems designed to support interpretable machine-to-machine interaction over a network. It has an interface described in a machine-processable format (specifically WSDL). Other systems interact with the web service in a manner prescribed by its description using SOAP messages, typically conveyed using HTTP with XML serialization in conjunction with other Web-related standards.'' When developers declare a new web service, it will be
discovered based on its description, which fully discloses its functionalities. Developers also have to declare a public interface and a readable documentation to help other developers when integrating different services \cite{w3cwsdl}. Nowadays, web API, standards that do not require XML-based web service protocols like SOAP and WSDL are also emerging. They are called REST (representational state transfer) services, which are moving towards simpler communication protocols.
%They are not restricted
%to XML formats, recently JSON, a human readable and simpler format
%is becoming popular among online service providers.

We are not going to delve into the engineering details of online web service implementation and its protocols in this paper. We are interested in web services from a business model perspective. Service providers usually charge end users for services they provide. For example, Google has listed pricing and plans for a wide range of services they provide on their web service console page\footnote{https://code.google.com/apis/console}.

In this paper, we abstract web services as rational agents\footnote{The term rational is used here in the sense that web services are utility maximizers} that provide services to end users. They aim to maximize their individual income  by receiving enough requests from end users. In order to increase their revenues, web services seek for more tasks if they have the capacity and throughput to do so. Web services can join communities to enhance efficiency by collaborating with others, to have access to broad market share, and for the opportunity to receive a bigger task pool from end users. Furthermore, the high reliance on web services has resulted in increased quality expectations from end users. Communities of web services can provide higher availability, performance, reliability, and recovery for end users.

\subsection{Web Service Communities}\label{s:wsc}

The community of web services is essentially a virtual group of web services having similar and complementary functionalities\cite{DBLP:journals/ijebr/MaamarSTBB09}. Communities aggregate web services and communicate with other entities such as UDDI registries and users, using identical protocols to those used by single web services. Web services join communities to increase utility by having a larger market share and task pool. Community coordinators are responsible for community development, managing membership requests from web services and distributing user tasks among the community members. Community coordinators try to attract quality web services to join and keep the community as stable and productive as possible to gain a better reputation and user satisfaction, which increases the community's market share. How web services reside within communities and how communities of web services are engineered is described comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}.

\subsection{The Join Challenge}\label{s:tjc}
Sometimes, web services can increase their overall utility by collaborating with other web services as communities. This collaboration helps them with better ways of sharing resources and having higher reputation, better market share and wider visibility. Web services and communities come with different internal and external parameters, and the long-term outcome is dependent on these parameters.

The goal of all parties involved in the community is to maximize their long-term outcome while they are operating as part of the community. Web services need to be equipped with a selection strategy to choose from all of the different possible collaboration groups they can form as well as an estimation method for evaluating the long-term gain of joining different possible communities. Web services need to experiment with different possible collaborative groups and other web services in order to estimate their gain over time. However, with a high number of possible communities, it is not possible to test collaboration with random web services. Even if a linear approximate function for estimating utility based on community web services' parameters is adopted, the exponential \footnote{Bell number: $http://en.wikipedia.org/wiki/Bell_number$} growth rate of the possible number of partitions of web services into communities would make any brute-force type algorithm for the best community selection strategy intractable and impractical in real-world application settings.

\subsection{Join Consequences}\label{s:jc}
It is worth mentioning that a \emph{join} event takes place as a result of two parties that are looking to expand their collaborations. All actions are chosen in an attempt to enhance the overall outcome. However, the selected action may result in decreasing the overall utility in the long run.
% long term.
This is the case when a single web service joins a community, but the complex process of task allocation eliminates the visibility of the web service and, thus, the web service stays idle within the community, which makes the join action a bad decision. The same event might be beneficial for the community, as it hosts a new web service that can engage in performing a new coming task. But overall, neither side benefits from collaborating with the other and the join event has negative consequences for at least one side's utility.

The more common scenario is when both parties benefit from the joining of a web service to a community. This action is rational, as both the web service and the community enhance the utility. However, the community may not be the best choice for the web service to join. In other words, the web service could have joined a better community, but, due to of lack of knowledge about the surrounding environment, joined the available community which was not the optimal choice. Since the community does enhance its utility, the web service could stay with that community, which results in a sub-optimal increase in web service's utility. In the following section the proposed model provides solutions that effectively address the aforementioned challenges.

\section{The Model Components}\label{s:themodelcomponents}

In this section, we discuss the parameters that we use in the rest of the paper. Then we present the task distribution and revenue model of our distributed web services communities.

\subsection{Internal Features}\label{s:if}

With a group of web services having identical or similar functionalities, QoS metrics provide nonfunctional characteristics for optimal candidate selection. Web service quality metrics were presented in \cite{Ardagna:2007:ASC:1263152.1263531,Menasce:2002:QIW:613357.613758,10.1109/ISSRE.2011.17}. We have adopted the most representative QoS properties of web services that highly influence the utility of web services.
%We refer to a typical web service as ws_{i}.

Let $C = \{ws_1,ws_2,..., ws_n\}$ be a community with $|n|$ web services. We define the following features for the group of web services based on their functional parameters:

\begin{itemize}

  \item \emph{Throughput $(Th_{C})$} is the rate at which a service can process requests. QoS measures can include the maximum throughput or a function that describes how throughput varies with load intensity. Throughput is a positive real number. For communities the expected throughput value of collaborating web services can be estimated as summation of throughputs of all the services inside:

	\begin{equation}
		 Th_{C} = \left\{
			\begin{array}{l l}
				Th_{w} & \quad \text{if $|C| = 1$}\\
				\sum_{w \in C}{(Th_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	\item \emph{Availability $(A_{C})$} is the percentage of time that a service is operating.
	The probability that the service operation is accessible. Availability is also a real number in the range $[0, 1]$. For communities the expected availability of collaborating web services considering they follow a parallel system can be estimated as:

	\begin{equation}
		A_{C} = \left\{
			\begin{array}{l l}
				A_{w} & \quad \text{if $|C| = 1$}\\
				1-\prod_{w \in C}{(1-A_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	\item \emph{Execution Time $(Et_{C})$} is the time a service takes to respond to various types of requests.
	%is the expected delay between the time instant when a request is sent and the time when the result is obtained.
	Execution time is measured in milliseconds. Execution time can be affected by load intensity, which can be measured in terms of arrival rates (such as requests per second) or number of concurrent requests. This internal feature is a positive integer. For communities, the expected execution time of collaborating web services can be estimated as the execution time of the bottleneck service which is the service with the slowest execution time:

	\begin{equation}
		Et_{C} = \left\{
			\begin{array}{l l}
				Et_{w} & \quad \text{if $|C| = 1$}\\
				max_{w \in C}{(Et_{w})} & \quad \text{if $|C| > 1$}
			\end{array} \right.
	\end{equation}

	%\item \emph{Data Quality} The ability of a data collection to meet user requirements , defined as the proximity of a value v returned by web service to a value considered as correct. The measure of data quality is considered here as a real number in the range [0, 1], where 1 represents the most desirable score.
\end{itemize}

	We normalize the range of these features so that each feature contributes approximately proportionately to the final utility outcome value. We adopt the \emph{standardization} method. We subtract the \emph{mean} from each feature then we divide the values of each feature by its \emph{standard deviation}.

%$HI = \overline{HI}$

\subsection{External Features}\label{s:ef}

The quantitative values of quality metrics need some benchmark values to represent their goodness. In other words, without some benchmark values it would be difficult for web services to identify their goals, and performance quality at any specific value of these metrics. Therefore, we have introduced two external features for assessing web services' estimate with regard to their standing among other web services.

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between a web service's \emph{execution time} metric and the maximum value of execution time of all the other web services, about which the web service has information of them. The less value means web service has better execution time compared to other web services.
	\begin{equation}\label{exp_1:f}
		exp_1 = Et_{i} - Et_{max}
	\end{equation}
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between a web service's \emph{throughput} metric and the maximum value of throughput of all the other web services,, about which the web service has information of them. The less value means web service has better throughput compared to other web services.
	\begin{equation}\label{exp_2Lf}
		exp_2 = Th_{max} - Th_{i}
	\end{equation}
\end{itemize}

A community of web services is abstracted as a standard web service that provides services that share the same internal and external features mentioned for web services. Inside, they adopt different protocols to manage the community; however, other web services or end users can communicate with communities of web services, using standard protocols, and inner details will be encapsulated for their clients.

\subsection{Task Distribution}

Communities of web services usually employ an implementation of Contract-Net protocol for task distribution, in which services opt for incoming tasks, and receive the tasks for which they ask. However, in our model, our community members would try to distribute tasks based on their capabilities and the QoS parameters provided by the web services. We have used a slightly modified \emph{weighted fair queuing} method to distribute tasks among community members. The goal is to allocate incoming tasks to web services with a rate matching the throughput value of $Th_{ws}$. In the \emph{weighted fair queuing} method, the input flow is multiplexed along different paths; however, in our case, if the rate of incoming tasks is less than the community's total throughput $(Th_{C})$, which is the summation of throughput values of the web services in the community, some of the input tasks will be queued and served with a delay.
%Thus, the amount of tasks performed by the community is $\sum_{ws}{Th_{ws}}$ when $\sum_{ws}{Th_{ws}} \leq R_{C}$.
However, when the incoming task rate is less than the throughput of the community, the \emph{weighted fair queuing} algorithm assigns a weighted task rate of $|input Task Rate| \times \frac{Th_{ws}}{\sum_{ws}{Th_{ws}}}$ for each web service ($ws$) within the community.

While distributing tasks, the community members can verify the performance, throughput and quality of service of tasks being performed by web services. The community can assess if those web services are capable of performing the number of tasks they advertised. If for any reason, there is a decline in the quality metric or throughput, the community can consider the new values as a benchmark for future performance calculations, and penalize the suspicious web services. This way, players will have incentive to truthfully disclose their actual capabilities in order to maximize profit from the community and to avoid being penalized. In addition, the system should be dynamic enough to detect and react to web services' quality metrics variation, as over time, web service metrics may degrade or improve, changes to which the community should adjust.
% Therefore its easy for the system to encourage players to be in some sense incentive compatible in the way that they would profit best by truthfully revealing their capabilities. Also it is important to be dynamic enough to consider web services which may have their quality metrics degraded or even improved over time for any reason and be able to adjust the community with new parameters.

\subsection{Community Revenue}

Communities and web services earn revenue by performing tasks. The total gain is a function of the quality of rate of tasks being performed. The utility of a collaborative group of services $U_{C}$ is the revenue of the community. Formula \ref{u_c_normal} is an estimation of the gain of the community.
